Live and Let Die? (Delayed Issues not Harder to Resolve)

### 2015-08-02 lucasRelatedWork.tex
Updates from ShareLaTeX

This was a common theme in the literature reviewed for this paper-- i.e.  that  it was no longer possible to access
the data used to make prior conclusions.
As an example of this, \fig{steck} shows one 2004 survey that reports an exponential delayed issue  effect for 
requirements issues in eight case studies. The right-hand-side of that column shows which of those referenced studies
we can still access-- and a ``?'' denotes a references from a broken url link (note that most
of those links are broken). Note that nearly all of that data is now no longer accessible.

@@

 \begin{figure}
{\small
\begin{center}
\begin{tabular}{r|rrrr|c}
 case& \multicolumn{4}{c|}{Phase Requirements Issue Found }&  \\
 study               &Requirements & Design & Code&  Test& source\\\hline
1 & 1& 6& 10& 70&~\cite{Boehm81}\\  
2& 1 &3& 5& 37&?\\ 
3&  1 &3& 7& 51& ?\\ 
4& 1& 5 &33 &75& ? \\
5& 1  &    & 10 & 40& ?\\ 
6& 1   &    & 10 &  40& ?\\
7&  1 & 20 & 45 & 250& ?\\  
8&  1  &   5 &      & 50& ? \\\hline
median           & 1 &  5  & 10   & 50.5 
\end{tabular}
\end{center}}
\caption{Sample of the Stecklein et al. literature review~\cite{steck04}:
cost to resolve requirements issues, relative to resolving issues during requirements.
Question marks ``?'' denotes a reference to
a result that is not longer accessible (i.e. a broken URL link).}\label{fig:steck}
\end{figure}

\begin{figure}[!b]
 \includegraphics[width=3in]{img/clutterbuck.png} 
 \caption{Cost of change from an agile case study. From~\cite{Clutterbuck09}.}\label{fig:clutterbuck}
 \end{figure}

 All in all, the empirical results of this section seem insufficient to justify
 the widely-held beliefs documented in the previous section. Adding to our doubts
 was the discovery of at least two studies that report less-than exponential
 increase in the difficulties associated with making the changes associated with delayed
 issues. For example,
Clutterbuck et al.~\cite{Clutterbuck09} studied 5-month effort by a small-to-medium enterprise team developing a 71KLOEC web interface to a database application to implement 18 change requests-- see Figure~\ref{fig:clutterbuck} (note that these were for new and changed user requirements, not defects). Clutterbuck et al. found the cost of change to be relatively flat until the later phases, with much of the effort spent in analysis of the change requests~\cite{Clutterbuck09}. Note that in this study, the effort increased by only 60\% (see the start and end of the curve in Figure~\ref{fig:clutterbuck}).

>>>>

This was a common theme in the literature reviewed for this paper-- i.e.  that  it was no longer possible to access
the data used to make prior conclusions.
As an example of this, \fig{steck} shows one 2004 survey that reports an 
exponential delayed issue  effect for 
requirements issues in eight case studies. Note that we cannot verify
{\em any} of those results since the links in the references of
that  survey are all broken (``page not foind'' errors).

@@

 \begin{figure}
{\small
\begin{center}
\begin{tabular}{r|rrrr}
 case& \multicolumn{4}{c}{Phase Requirements Issue Found }  \\
 study               &Requirements & Design & Code&  Test\\\hline
1& 1 &3& 5& 37\\  
2& 1  &    & 10 & 40 \\ 
3& 1   &    & 10 &  40 \\
4&  1  &   5 &      & 50  \\
5&  1 &3& 7& 51 \\
6& 1& 5 &33 &75  \\
7&  1 & 20 & 45 & 250 \\  
\end{tabular}
\end{center}}
\caption{Sample of the Stecklein et al. literature review~\cite{steck04}:
cost to resolve requirements issues, relative to resolving  during requirements.}\label{fig:steck}
\end{figure}

\begin{figure}[!t]
 \includegraphics[width=3in]{img/clutterbuck.png} 
 \caption{Cost of change from an agile case study. From~\cite{Clutterbuck09}.}\label{fig:clutterbuck}
 \end{figure}

 All in all, the empirical results of this section seem insufficient to justify
 the  beliefs documented in the previous section. Adding to our doubts
 was the discovery of   several studies that report less-than exponential
 increase in the difficulties associated with making the changes associated with delayed
 issues. For example,
Clutterbuck et al.~\cite{Clutterbuck09} studied 5-month effort by a small-to-medium enterprise team developing a 71KLOEC web interface to a database application to implement 18 change requests-- see Figure~\ref{fig:clutterbuck} (note that these were for new and changed user requirements, not defects). Clutterbuck et al. found the cost of change to be relatively flat until the later phases, with much of the effort spent in analysis of the change requests~\cite{Clutterbuck09}. Note that in this study, the effort increased by only 60\% (see the start and end of the curve in Figure~\ref{fig:clutterbuck}).

 Other examples, where research papers report non-exponential changes in the difficuty in resoloving issues
are:
  \bi
  \item
  Boehm~\cite{Boehm80} reported a flatter growth rate for small, non-critical projects.
   \item 
  Data from NASA's Johnson Space Flight Center, reported by Shull~\cite{Shull02}, found that the cost to find certain non-critical classes of defects was fairly constant across lifecycle phases (1.2 hours on average early in the project, versus 1.5 hours late in the project).

\ei

### 2015-08-02 phaseDelay.tex
Updates from ShareLaTeX

One technical aside: there are several ways to  measure {\em difficulty}. The first is a direct measure of the actual work time spent in correcting the defect in the product. Direct measurement enables analysis of the variability the cost-to-fix among a category of defects. The second approach measures the total time spent in a defect removal phase and divides by the total defects removed. This provides an average cost-to-fix including overhead of that activity. For example, a code inspection may require an hour of effort to remove three defects for a 3 defects per hour or to invert the ratio, 1/3 hours per defect. One observation made by Nichols while studying our 171 projects is that most of the time  is spent locating the defects not actually fixing them. we use the second approach in our analysis.

@@

\begin{figure}[!t] 
%\vspace{-1cm}
%\hspace{-0.35in}
\begin{center}
\includegraphics[width=3in]{img/boehm-overlay.pdf}
 \end{center}
% ~\\~\\
%~\\~\\~\\

 \caption{Cost-to-fix curve from \protect\fig{cost-to-fix} 
 overlayed with case study from~\cite{Royce98} and TSP data. Scale is relative to the phase with the lowest cost to fix. }\label{fig:cost-to-fix-tsp}
 \end{figure}

@@

Using Figure~\ref{fig:fix-phase-dist}, we can comment on two specific and more general
delayed issued effects. 
\bi
\item {\bf Specific1}: Requirements errors are the most expensive to fix;
\item {\bf Specific2}: Requirements errors are the most expensive to fix, 
 when found during production but the cheapest to fix early in development;
 \item {\bf General:}  It is exponentially more  difficult  to resolve an issue in a software project, the   longer it remains.
\ei
\subsubsection{Specific1}

Our data does not support   
{\em  requirements errors are the most expensive to fix}.
Note that   median time spent fixing errors from {\em requirements} across the life cycle was 92, 93,  49, 33, 39, 49, 67, 122 minutes\footnote{If this time seems long, then recall
that it includes the time required  to (a)~collect data and realize there is an error;
(b)~prepare a fix;  and (c)~apply some validation
procedure to check the fix (e.g. discuss it with a colleague or execute some tests).}.
This is:
\bi
\item
Clearly longer than median times seen {\em Before } and in {\em Planning}
(30,28,45,77,65,56,24,35,33);
\item
Somewhat shorter, but still not exponentially smaller,  than the median times see after {\em requirements} in {\em design}
(103,101, 79,66, 119,88, 80,45) and in {\em code} (65,115, 113,125, 95,133). 
\ei
For completeness, we ran an  
A12 test comparing the raw data (and not just the medians of \fig{raw}) and found
 a non-small effect only for the difference {\em requirements}
to {\em code}. But even though the effect was non-small, 
 the overall size of the difference
{\em requirements} to {\em code} was less than 30\% (i.e. much less than we would
expect from the delayed issue effect). 
\subsubsection{Specific2}
Next, we comment on 
requirements errors are the most expensive to fix 
{\em when found during production but the cheapest to fix early in development}}. 
Our data from these 171 project does not extend into post-delivery deployment. That said,
there is enough evidence to raise doubts that our projects will exhibit exponentially more
difficult issues after deployment. 
Specifically: prior reports of an  exponential increase in difficulty in resolving issues {\em during
deployment} might have been used to predict for exponentially increasing
difficulty {\em post-deployment}.
For example:
\bi
\item  From requirements to acceptance testing
\fig{b81} reports a 40-fold increase;
\item
 \fig{cost-to-fix} reports a 100-fold increase
(for the larger projects).
\item
The projects surveyed in \fig{steck} report changes of a  \newline
\{37,40,40,50,51,70,75,250\}-fold increase in pre-deployment period when the software was being developed.
\ei
On the other hand, a remarkable feature of Figure~\ref{fig:fix-phase-dist} is the {\em absence}
of an exponential delay effect in the difficulty to resolve requirements issues during development.
The greatest difficulty in resolving requirements issues was 1.31 seen in systems testing--
which is far smaller than the 37 to 250-fold increases reported in \fig{b81}, \fig{cost-to-fix},
and \fig{steck}.  That is, from our TSP data, we would not anticipate exponentially greater difficulty in resolving requirements issues after
deployment. 
 \subsubsection{General}
 The above specific statements related just to requirements. Turning now
 to issues creased in all phases in Figure~\ref{fig:fix-phase-dist},  we see that
 this data does not follow the exponential increase predicted
by the delayed issue effect:
\bi
\item There is much ``up and down'' in the chart. 
\item If we separate out just the {\em Inspect} or {\em Review} data, we can see that there is some increase in {\em Design} and {\em Code}. However, that effect could be a linear or be just noise. In either case,
it does not look like exponential.
\item The time in final testing is particularly low suggesting that few defects survived into testing
(and hence, there were few instances where code required massive rework).
\ei

For illustrative purposes, the relative cost to fix errors per lifecycle phase in the TSP project data is overlayed with Boehm's cost curve~\cite{Boehm81} as well as the CCPDS-R case study~\cite{Royce98} in \fig{cost-to-fix-tsp}. Cost-to-fix for TSP is total time spent in each review, inspection, or test-related phase / \# of defects found in phase (reported in \fig{raw}). It is clear to see that the trend is relatively flat, closer to the pattern exhibited by the Royce case study (which also focused on early defect removal) than the exponential growth curve.

@@

\section{Validity}

To sum up the paper so far,
\bi
\item
The delayed issue effect has been reported in around a dozen papers, at least half of which date back to the 1970s. 
\item
But there exists
no evidence for this effect was seen  in hundreds of software projects build   since 2006.
\ei
From this second point, is it valid to assert that it is {\em not} disruptive to  software projects
to let some issues advance from one phase to the next?

\subsection{Alternate Explanations}

One explanation for the above is that our sample is biased-- that while general software
systems suffer from the delayed issue effect, systems built using TSP are so much better controlled
that they  mitigate against the DIE effect.

It would be convenient to offer the last paragraph as the conclusion of this study. Two of our authors (Nichols and Shull)
 work at the organization that invented and promoted TSP.

\subsubsection*{H4: A post-development DIE effect can be predicted from pre-deployment data} 
The data collected from our 171 projects     pertains only to
development, and not to subsequent operations.
So did we miss the DIE effect in our data since
our data collection is just not prolonged enough to detect that effect?

We would argue that this is not the case.
An  exponential DIE effect can be seen {\em during the development} of all
the papers we have reviewed. For example:
\bi
\item In \fig{b81}: see the data relating to development and acceptance testing;
\item In  all the case studies of \fig{steck} in the columns {\em Requirements} to {\em Test}.
\ei
That is, in the literature we have surveyed, the post-development DIE effect would have been predicted by data collected during pre-deployment development.
Such

Another surprise was a lack of {\em growth trend} suggesting that, in operation,
 we should expect larger exponential delays in resolving issues. This is odd since all
 the papers we could access on DIE should such a growth trend in the pre-operational data.

\subsubsection*{H5: ``Time'' to resolve an issue is analogous to ``cost''}
This hypothesis a somewhat technical point needed for our argument.
While classical papers on the delayed issue effect discuss the {\em cost}
of delaying issues, the data from our 171 projects only has time data.
Hence, to say that our data addresses the classic delayed issue effect,
we need to make the case that our time measurements are an appropriate
surrogate for the prior claims on the cost of delaying issues.

We report validity concerns following the guidelines of Runeson \& H\"{o}st~\cite{runeson09}.

\subsection{Construct validity}
XXXX die does not aoccur during developmentjust during deployment

that means during period requirements to test pre-deployent, there is not rush to resolve issues (just as long as a concerted effort is made to retire all issues priot to deployment)

all our data is skewed due to TSP. two responses. 1) we don't thing so. 2) sure, let it due to TSP in which case we can fix the delayed issue effect via process methods 9everyone adopts the team software process). In either case, the delayed issue effect can be ignored or mitigated without extensive investment in radically new methods or tools (e.g. automatic formal methods, better requirement description languages).

The essential constructs in our analyses are \emph{cost-to-fix, defects, and phase delay}.

\emph{Cost-to-fix} is traditionally measured as either effort spent to fix a defect or time spent fixing a defect. In our analysis, we calculate the latter as \# of defects fixed in a defect-removal plan item, such as DesignReview or CodeInspection, divided by the time spent in the plan item. A single project has multiple such plan items. In the TSP data, this time accounts for time spent isolating, analyzing, and fixing the defect, but does not include time spent verifying the fix.

In our sample, all \emph{defects} are considered equally, and thus TSP's reporting methods may result in ``small'' defects could be dominating the sample. For example, typographical requirements errors are included as well as requirements errors that necessitate an architectural change to the system. Our analysis does not include a notion of defect severity, however, we observe no indications in previous cost-to-fix results reported by \cite{Boehm81} and \cite{Royce98} that severity was considered in their samples. However, evidence discussed in \cite{Shull02} suggests that low severity defects may exhibit a lower cost to change. Nonetheless, even ``small'' typographical errors have been known to cause enormous damage (e.g., the Mars climate Orbiter).

\emph{Phase delay}, as depicted in the scale up column in \fig{cost-to-fix-tsp} and \fig{raw}, is a discrete value that does not consider the total time that a defect is latent in the system. Thus, the phase delay of two defects injected in Code but found in System Test will be the same even though one was found at the beginning of test while the other at the end of a potentially lengthy test period. Ideally, one could examine the total real time a defect was in a system from the point at which is was injected to the point at which it was discovered. It may be possible to gather such information if the project keeps rigorous change logs for its requirements, design, and code artifacts.

@@

\subsection{External validity}
Threats to external validity affect the generalization of findings. The most obvious threat to external validity is that TSP is not representative of most software development processes. TSP was, in fact, designed to be rigorous and eliminate defects early in the lifecycle and soon after injection. We make no assertions regarding the generalizability of these results beyond TSP, we observe that other case studies in various domains have exhibited flattened cost-to-fix curves.

Our analysis of cost-to-fix does not include defects found in production, which, in Boehm's original data, account for costs up to 200 times the cost of finding and fixing defects in Requirements. Further, it is not clear how Boehm~\cite{Boehm81} and Royce~\cite{Royce98} define \emph{cost-to-fix}. In our analysis, cost-to-fix includes the time spent analyzing, fixing the defect, and validating the fix. If Boehm and Royce do not include all of these costs, then the TSP cost curve in \fig{cost-to-fix-tsp} artificially high by comparison.

@@

In this paper, we have examined contemporary project data for evidence to support the well-known software engineering truism of the phase delay effect. 
We found that, in the literature, the evidence for phase delay was very spare.
%and (in a $21^{st}$).  %don't know what this is supposed to be
We suspect that a very old result was requoted, without being 
critically re-examined, for literally decades.

@@

One hypothesized explanation that can be ruled out is that this lack of
a phase delay effect is due to the use of TSP.
  To the contrary, four non-TSP  case studies reported that the expected exponential growth in cost-to-fix was not observed:
  \bi
  \item
  Boehm~\cite{Boehm80} reported a flatter growth rate for small, non-critical projects.
  \item
  Clutterbuck~\cite{Clutterbuck09} showed a flatter cost curve in an agile case study. 
  \item 
  Data from NASA's Johnson Space Flight Center, reported by Shull~\cite{Shull02}, found that the cost to find certain non-critical classes of defects was fairly constant across lifecycle phases (1.2 hours on average early in the project, versus 1.5 hours late in the project). 
  \item
  Royce~\cite{Royce98} reported a flat cost curve for a project that invested heavily upfront in architectural reviews.
\ei
If the lack of phase delay is not due to TSP, then what does cause it?
We surmise that the substantial changes in software development and deployment technologies and practices, since the phase delay effect was first measured, may be responsible for our observations in the TSP data set. Recall that the original data on which the phase delay effect is based, as reported in \S3, are approximately 40 years old. The most recently reported data, from Shull et al. \cite{Shull02}, was published in 2002 but contains substantial retrospective data. Over time, it would be surprising if the investment in better languages, IDEs, and other development technologies had in fact not created some improvement. To pick just one example, we note that much of the data showing the phase delay effect was collected before frequent and unobtrusive auto-updating software was as common as today, which certainly flattens the cost of changes.

@@

Changes to  software that is used by a large population can be confusing (particularly for the sick or elderly). 
Consider the 23,000,000 veterans working with  the U.S.  Veterans 
Administration (VA) service. Some of those veterans are dependent on health care or pensions from the VA. Those veterans use the VA's web interfaces
to  (e.g.) find out
when their next cheque will arrive. Before we can update the VA web software, we would
need to  a transition plan that helps users   move from the older to the newer system. This is an example of what might be called a ``user phase delay''; i.e. changes to software are more complex to process, particularly if some users have styled
their day-to-day work flow around the services offered by the older software.

>>>>

@@

@@

Using Figure~\ref{fig:fix-phase-dist}, we can comment on two specific and more general
delayed issued effects. 
\bi
\item {\bf Specific1}: Requirements errors are the most expensive to fix;
\item {\bf Specific2}: Requirements errors are the most expensive to fix, 
 when found during production but the cheapest to fix early in development;
 \item {\bf General:}  It is exponentially more  difficult  to resolve an issue in a software project, the   longer it remains.
\ei

\subsubsection{Specific1}

Our data does not support   
{\em  requirements errors are the most expensive to fix}.
Note that   median time spent fixing errors from {\em requirements} across the life cycle was 92, 93,  49, 33, 39, 49, 67, 122 minutes\footnote{If this time seems long, then recall
that it includes the time required  to (a)~collect data and realize there is an error;
(b)~prepare a fix;  and (c)~apply some validation
procedure to check the fix (e.g. discuss it with a colleague or execute some tests).}.
This is:
\bi
\item
Clearly longer than median times seen {\em Before } and in {\em Planning}
(30,28,45,77,65,56,24,35,33);
\item
Somewhat shorter, but still not exponentially smaller,  than the median times see after {\em requirements} in {\em design}
(103,101, 79,66, 119,88, 80,45) and in {\em code} (65,115, 113,125, 95,133). 
\ei
For completeness, we ran an  
A12 test comparing the raw data (and not just the medians of \fig{raw}) and found
 a non-small effect only for the difference {\em requirements}
to {\em code}. But even though the effect was non-small, 
 the overall size of the difference
{\em requirements} to {\em code} was less than 30\% (i.e. much less than we would
expect from the delayed issue effect). 
\subsubsection{Specific2}
Next, we comment on 
requirements errors are the most expensive to fix 
{\em when found during production but the cheapest to fix early in development}. 
Our data from these 171 project does not extend into post-delivery deployment. That said,
there is enough evidence to raise doubts that our projects will exhibit exponentially more
difficult issues after deployment. 
Specifically: prior reports of an  exponential increase in difficulty in resolving issues {\em during
deployment}.
For example:
\bi
\item  From requirements to acceptance testing
\fig{b81} reports a 40-fold increase;
\item
 \fig{cost-to-fix} reports a 100-fold increase
(for the larger projects).
\item
The projects surveyed in \fig{steck} report changes of a  \newline
\{37,40,40,50,51,70,75,250\}-fold increase in pre-deployment period when the software was being developed.
\ei
Any manager watching  these intra-development trends would be well-justified
to anticipate that  the difficulty in resolving issues would will get much worse. Such managers
would therefore expect an exponentially harder time resolving issues during deployment.

On the other hand, a remarkable feature of Figure~\ref{fig:fix-phase-dist} is the {\em absence}
of an exponential delay effect during development.
In our data, the greatest difficulty in resolving requirements issues was 1.31 seen in systems testing--
which is far smaller than the 37 to 250-fold increases reported in \fig{b81}, \fig{cost-to-fix},
and \fig{steck}.  That is, from our  data, we would not anticipate exponentially greater difficulty in resolving requirements issues after
deployment. 
 \subsubsection{General}
 The above specific statements related just to requirements. Turning now
 to issues creased in all phases in Figure~\ref{fig:fix-phase-dist},  we see that
 this data does not follow the exponential increase predicted
by the delayed issue effect:
\bi
\item There is much ``up and down'' in the chart. 
\item If we separate out just the {\em Inspect} or {\em Review} data, we can see that there is some increase in {\em Design} and {\em Code}. However, that effect could be a linear or be just noise. In either case,
it does not look like exponential.
\item The time in final testing is particularly low suggesting that few defects survived into testing
(and hence, there were few instances where code required massive rework).
\ei
In summary:
\bi
\item
The delayed issue effect has been reported in around a dozen papers, at least half of which date back to the 1970s. 
\item
But there is no trace of that  effect   in hundreds of software projects built  since 2006.
\ei
%For illustrative purposes, the relative cost to fix errors per lifecycle phase in the TSP project data is overlayed with Boehm's cost curve~\cite{Boehm81} as well as the CCPDS-R case study~\cite{Royce98} in \fig{cost-to-fix-tsp}. Cost-to-fix for TSP is total time spent in each review, inspection, or test-related phase / \# of defects found in phase (reported in \fig{raw}). It is clear to see that the trend is relatively flat, closer to the pattern exhibited by the Royce case study (which also focused on early defect removal) than the exponential growth curve.

%\begin{figure}[!t] 
%%\vspace{-1cm}
%\hspace{-0.35in}
%\begin{center}
%\includegraphics[width=3in]{img/boehm-overlay.pdf}
 %\end{center}
% ~\\~\\
%~\\~\\~\\

 %\caption{Cost-to-fix curve from \protect\fig{cost-to-fix} 
 %overlayed with case study from~\cite{Royce98} and TSP data. Scale is relative to the phase with the lowest %c%ost to fix. }\label{fig:cost-to-fix-tsp}
% \end{figure}

@@

\section{Validity}
We report validity concerns following the guidelines of Runeson \& H\"{o}st~\cite{runeson09}.

\subsection{External validity}
Threats to external validity affect the generalization of findings. The most obvious threat to external validity is that TSP is not representative of most software development processes. Hence,
For example, 
one way to discount the above is that our sample is biased-- that while general software
systems suffer from the delayed issue effect, systems built using TSP are so much better controlled
that they  mitigate against the DIE effect.

It would be convenient to offer the last paragraph as the conclusion of this study
(two of our authors, Nichols and Shull,  work at the organization that invented and promoted TSP).
 However, we cannot endorse such a conclusion:
 \bi
 \item
We view TSP as a better way to
 {\em monitor} the activities {\em of  existing projects}. That is, TSP
 does not radically change a project, it just offers a better way to log the activity within
 that project. 
 \item
 Our 171 project contain examples of a wide variety
 of systems (e-commerce web portals and  banking systems) run in a variety of
 ways (agile or  waterfall or some combination of the two). Hence, rather
 than being unduly  biased in some manner,
 we view our 171 projects as a large sample of current industrial processes for software
 development.
 \ei
Another way to discount this result is to say we are only reporting on small to medium-size
projects, and the real impact of delayed issues is on very large projects.
There certainly exists another class of very large software project where early requirements
errors would be catastrophic for later development. For example, when NASA launches deep space missions,
each of those satellites are special-purpose, uniquely-built devices built by a large contractor
population. Such systems are very hard to reconfigure halfway through the development.

We concede that for very large complex NASA-like developments, the delayed issue
effect could still hold. However, perhaps we need to split our software engineering principles
into two parts- one for   very large NASA-like projects (which are not very numerous)
and one for smaller-sized projects like the ones studied here (which are far more numerous).
Most software is {\em not} like the software used on NASA deep-space satellites.
As  agile methods grew more popular last decade; and as personnel computers grew more powerful;
and as the libraries associated with particular languages grew more extensive; then smaller
teams starting  deliver systems that (previously) had required very large teams.
As mentioned above,
our  171 projects represents $\frac{171}{212}=80$\% of the software projects
seen by the Software Engineering Institute. That is, as far as we can tell,
there are more small to medium-sized projects than otherwise.

\subsection{Construct validity}

Our definition of {\em difficult to resolve} combines two concepts: time to change and cost to
change. In the above we have used them interchangable, comparing our time to resolve data (from the 171 TSP projects) against the cost-to-fix results of Boehm et al. e.g. \fig{b81}.
Is it valid to assume the equivalence of time and cost?

Certainly, there are cases where time is not the same as cost. Consider, for example,
if debugging required some very expensive tool or the services or a very senior (and hence, very expensive)
developer. Under those circumstances, time does not equate to cost.

Also, there is the issue of the implication of  a defect. Evidence discussed in \cite{Shull02} suggests that low severity defects may exhibit a lower cost to change. Nonetheless, even ``small'' typographical errors have been known to cause enormous damage (e.g., the Mars climate Orbiter).

Having documented the above issues, we assert that they are probably not major issues in the study.
As to the first point, that some debugging may be more expensive than others, one of us (Nichols)
was closely associated with many of the projects in our sample. He is unaware of any frequent
use of exorbitantly expensive tools or people on these projects. As to the second point (that
some defects have an impact beyond the mere cost to fix them), we no evidence suggesting
that the net effect of this concern over all our projects is non-zero.

@@

@@

In this paper, we have examined contemporary project data for traces of
the widely-believed delayed issue effect.
We found that, in the literature, the evidence for this effect was very spare--
in fact, we found
%and (in a $21^{st}$).  %don't know what this is supposed to be
We suspect that a very old result was requoted, without being 
critically re-examined, for literally decades.

@@

One hypothesized explanation that can be ruled out is that this lack of
a phase delay effect is due to the use of TSP.
  To the contrary, four non-TSP  case studies reported that the expected exponential growth in cost-to-fix was not observed:

If the lack of phase delay is not due to TSP, then what does cause it?
We surmise that the substantial changes in software development and deployment technologies and practices, since the phase delay effect was first measured, may be responsible for our observations in the TSP data set. Recall that the original data on which the phase delay effect is based, as reported in \S3, are approximately 40 years old. The most recently reported data, from Shull et al. \cite{Shull02}, was published in 2002 but contains substantial retrospective data. Over time, it would be surprising if the investment in better languages, IDEs, and other development technologies had in fact not created some improvement. To pick just one example, we note that much of the data showing the phase delay effect was collected before frequent and unobtrusive auto-updating software was as common as today, which certainly flattens the cost of changes.

@@

Changes to  software that is used by a large population can be confusing (particularly for the sick or elderly). 
Consider the 23 million veterans working with  the U.S.  Veterans 
Administration (VA) service. Some of those veterans are dependent on health care or pensions from the VA. Those veterans use the VA's web interfaces
to  (e.g.) find out
when their next cheque will arrive. Before we can update the VA web software, we would
need to  a transition plan that helps users   move from the older to the newer system. This is an example of what might be called a ``user phase delay''; i.e. changes to software are more complex to process, particularly if some users have styled
their day-to-day work flow around the services offered by the older software.
